### 2015-10-10 doc/src/book/referee-reports/response.do.txt
Added source files.

TITLE: Reponse to referee reports on ``Finite Difference Computing with Exponential Decay Models''
AUTHOR: Hans Petter Langtangen
DATE: Today

======= Background =======

The original preface to this book explained the background for the book,
but this part was removed since I thought it was irrelevant for the
audience that shows enough interest to pick up the book. However, for
referees and editors the background can certainly be of interest as it
provides a response to some of the criticism in the reports.

The book started out as compact a 15-page appendix to define the
expected level in numerical methods and Python implementation required
for a course/book on finite difference and element methods for PDEs.
The course is at the MSc/PhD level at the University of Oslo and
attracts about 50 students with widely different backgrounds ever
year.  The students come from physics, mathematics, numerical
analysis, computer science, geoscience, life science, chemistry,
nanotechnology, and so on, with bachelor degrees from various types of
universities and university colleges in Norway and abroad.  The goal
of the course is to equip the students with all the *practical tools*
they need to do solve PDEs in their thesis work in a reliable,
reproducible, and human-efficient way. This means that any project in
the course involves the entire scientific computing pipeline, but with
a particular focus on programs that can be ``proved'' to work. Because
we emphasize the entire pipeline (and not just the mathematics, as was
usual in our previous traditional numerical analysis courses) there
are many different types of scientific topics, so most students will
be very familiar with some elements and struggle with others. In the
end, we produce a quite homogeneous group of ``PDE
practitioners''. The course is very popular with excellent student
critiques, and lots of supervisors who send their students to this
course and regard it as a ``gold standard'' on how to do PDE
computing. My main argument for sending this tutorial to Springer, is
that it works well to define a common knowledge ground in this course,
and the students like the text.

There is a strong tradition that courses on numerical solution of PDEs
are given by numerical analysis or scientific computing groups and
adapted to students specializing in these subjects.  A nice by-product
is that such courses often work well for physicists and mathematicians
too. Nevertheless, computing with PDEs grows rapidly throuhout the
sciences these days, and the computing technology must be taught to
much broader and heterogeneous audiences than what we are used to. I
believe the course I have designed in Oslo is a first step in this
direction and therefore provide relevant experience for the future
demands to teaching material.

The typical feature of such heterogeneous audiences is that some
students are very strong in mathematics, but surprisingly weak in
programming, and vice versa, while others have their strengths in
modeling and keeping track of all details in real-world problems. My
solution to this challenge of heterogeneity is to provide teaching
material where *all details are written out and explained*. Obviously,
lots of such detailed material will be boring for a given student, but
other parts are new and unfamiliar.  If new material is easy to read
and tightly coupled to computer implementations, my experience is that
students start doing computational experiments and spend hours on
exploring the subject and digesting the details.
On the other hand, if new material is described
in a compact way, with many mathematical and implementational details
left to the reader, the chances are high of losing interest, stopping up, and
not putting in hours of experimentation and deep learning.
This is why I have adopted over the years a writing style that is very
verbose and explores all ``nuts and bolts'' necessary to solve the
problems, with a tight coupling of mathematics and implementation.
Students like this style, but it goes without saying that
experts, like referees and colleagues, find the familiar material tedious and
boring, while the subjects they are not familiar with, typically
implementation, are often not considered so important or interesting. In this
landscape, I listen to the students and stick to the philosophy that
implementational aspects are at least as important as mathematical topics.

Back to the initial, brief appendix that led to this book: it listed
the required knowledge through a solution of the problem $u'=-au$, but
failed to *teach* any of this knowledge. I was asked to make the
example less brief, and this led to a series of requests for more
details, especially regardig implementational aspects.  The appendix
became a chapter with the software engineering topics fully integrated
throughout the text, then the software engineering parts were factored out
in a separate chapter, and finally all this material reached a size
that made it look odd as the first half of a book on PDEs.  The
mathematics and computer science level of this introductory material
is also much lower than for the PDE topics. In the course I described
above, the material is meant for self-study. It therefore made more
sense to factor out such introductory material in a separate tutorial.

Although I personally have used the tutorial only for refreshing known
subjects and supplementing lacking knowledge for students at the
MSc/PhD level, it is clear that it could be used very early
on in a bachelor program as a first taste of the scientific computing
pipeline for a broad audience. Only some elementary calculus and
computer programming is needed as background. The tradition is to
wait with ODEs until the mathematical foundation is good, but this
is neither motivating nor necessary. Whether the text would
function in, e.g., a second-semester course is impossible to say without
testing. All I know, is that over 150 students have used the material and
requested all the changes that made a compact appendix grow into the
proposed book.

As I comment upon below, Project 4.15 has been a compulsory starting
project to ``set the stage'' in my course, and the present tutorial
emerged partly as a response to the low performance in such a real-world
project. It was clear that the students needed more knowledge about
making reliable computer programs, even in mathematically very simple
cases.

======= Report 1 =======

=== Referee ===

Here is my general impression upon a quick reading.

It is always hard to write a critical assessment of a book proposal:
the author must have written a lot and invested a lot of effort and
time, just in order to enable the referee to have a chance of saying
something worthwhile. Here, fortunately, there is a proposal to cut
part of a planned book (one of three planned books[^other-books])
and make it into a
separate tutorial. So, even if the referee does not like the proposal,
there is a way out.

[^other-books]: Author's notification:
The tutorial was part of another book on finite difference
methods for PDEs. The two other books mentioned are about different topics
and not related to this tutorial.

And indeed I must confess that I am not thrilled with the packaging
proposed here. It is far from clear that a student learning ODE
methods must know the details of Python, let alone software
engineering, and it is far from clear that a software engineer must
know details of numerical methods for ODEs. Moreover, there are really
many accessible ODE texts, including Griffiths-David-Higham and
Ascher-Petzold, and also long chapters in some introductory numerical
textbooks, which are better written than this exposition.

=== Response ===

It is not difficult agree with the comments in the last paragraph. However,
the text assumes that the student's goal is the learn ODE methods
*and* make reliable programs for computing the solution. The goal is
not one of the topics, but the combination, because that is what is
required when doing real-world science or engineering.

No student of software engineering will every read this tutorial as it
is far too trivial and incomplete. Better sources of information
exists for those who can afford to spend the time reading them. Similarly, students
in numerical analysis or scientific computing will take several
numerical methods courses, most likely one devoted to ODE methods,
e.g., using the two mentioned excellent textbooks. This tutorial was
written for all those who follow other programs where there is no
space or interest for separate courses on ODE methods or software engineering, yet
the students have a demand for computing with differential equations
in a thesis context. Or the tutorial could be a very first taste of
these topics and their interplay very early on in the studies.

=== Referee ===

For instance, why are the Python tutorial and the ODE introduction in
Chapter 1 so tightly coupled? The Python tutorial can be given in an
appendix for the author's planned book, and for a Python example it
would suffice to consider forward Euler for $u'=-u$, which could be
introduced in one page (at most).

=== Response ===

The referee's suggestion is the normal way to do it in the
existing literature: compact
description of a method and a *separate* brief listing of programming aspects.
My experience is that these are not separate subjects, but tightly
coupled. When presented and taught separately, students struggle with the
understanding of the interplay between the topics. Methods, implementation,
and testing are equally important in a science project, and especially
testing is tightly coupled to the results of mathematical analysis.

It is true that the forward Euler method can be introduced in one page at most,
but to avoid that the method is memorized just as a formula, rather than being
well understood, it is key to spend time on the *ideas* and to show
how the same reasoning can produce alternative schemes. That is the rationale for
all the space used on introducing the thinking around finite difference
methods. I want the ideas to get through, not the formulas, and Chapter 1
is the convergence of this writing journey.

=== Referee ===

I believe that the student should first be motivated to look at the
simple ODE of Chapter 1 (referring to Chapter 4 does not do it),
before so many details are exposed in the first two chapters.

=== Response ===

I did not understand this comment, but would of course like to improve
the motivation in Chapter 1. Chapter 4 is factored out because the
material brings in so many topics from applied sciences that are
irrelevant for producing reliable programs.  Parts of Chapter 4 can be
used as motivation for the model in Chapter 1 (I always do that), but
a fundamental idea (as stated) is to always start with a stripped
mathematical model that captures just the essence of real-world
models. I chose $u'=-au$ over $u'=-u$ because the former has a
parameter, and this parameter will always be attached to the
discretization parameter as the key dimensionless parameter that
governs the quality of the numerical solution.

=== Referee ===

Note also that you need $a>0$ to talk about decay (are you
really?!).

=== Response ===

We do not need $a>0$, but the features of the numerical solution, especially
artifacts, are different for $a<0$ and left as an exercise to investigate.
I realize that $a>0$ was not mentioned in the opening of the text, but now
it is stated that $a>0$ is required for decay.

=== Referee ===

Further, the theta-method is far less fundamental than
forward Euler, bakward Euler, trapezoidal and midpoint.

=== Response ===

That is true, from a mathematical point of view, but the theta-method is
fundamental for implementation.  Without it, we get if-tests with
repetitive code for the three schemes.

=== Referee ===

Also, must an elementary student or a casual reader know in detail
about odespy, which is just another initial-value ODE package?

=== Response ===

There is a separate tutorial on using Odespy (available on URL:
"https://github.com/hplgit/odespy"), so the text in this tutorial is
meant to be brief. We show 1) a minimalistic example, and 2) how to
set up an experiment with lots of solvers. This information covers
about a page and is useful when doing many of the exercises in
Chapters 3 and 4.  The rest of the text seemingly related to Odespy
is about numerical performance
of different methods.

There is, however, a detailed remark about a less pedagogial aspect of
using implicit methods in Odespy. Motivated by the refeee's comment,
this remark is now moved to the code and left out of the tutorial.

=== Referee ===

If the author is aiming at PDEs then I don't understand the
essentially-nonexistent treatment of nonlinear problems when using
implicit methods. But that does lead again to the suggestion of a
different organization of the material.

=== Response ===

This is absolutely a valid comment. Methods for nonlinear problems
could well be introduced in Chapter 3. The reason for not doing this
is two-fold: 1) to limit the amount of mathematical topics in the
tutorial, and 2) because I have written an extensive chapter elsewhere
on nonlinear problems for ODEs and PDEs, and it appeared very natural
to do both ODEs and PDEs together in that chapter.

======= Report 2 =======

I have read quickly through Hans Petter Langtangen’s text and, to be
frank, I am undecided. To put it frankly, I hate the first part of the
book and love the second.

The first part of the book essentially assumes zero knowledge of
mathematics. But what is the audience?  Each and every undergraduate
student in physical sciences and engineering in any half-respectable
university has under their belt considerably more mathematical
knowledge than assumed in the first part of the text. This is not
``gentle'', it is so soft as to be utterly boring and repetitive,
exploring the same kind of very basic ideas first for a scalar linear
equation with a contact coefficient and no inhomogeneous term, then
for a scalar linear equation with a variable coefficient and an
inhomogeneous term and finally for systems of equations. Presumably,
students, before they even open the first page of such a book, have
some basic common knowledge. No matter how low the common denominator,
it will include vectors and matrices, basic definition and properties
of ODEs and, most importantly, basic understanding of how mathematics
works at its simplest.

=== Response ===

The referee's comment is certainly true for many students in
mathematically-oriented programs, but the new coming audiences from
geoscience and bioscience are significant in size, and these
students have a much weaker mathematical knowledge, understanding,
and interest,
even those who have taken a couple of mathematics courses.  My
experience is that if these students are given the present ``utterly
boring and repetitive'' text, rather than just pointing them to thick
math books to upgrade their mathematical foundation, they get enough
interest and drive to catch up, and they actually posess sufficient
other types of knowledge to complete a practially oriented PDE
computing course with a decent grade.

I should add that there have been many requests from students to
extend explainations in the present tutorial, but actually no comments
on making the text more compact (and this includes detailed student critiques
from 50+ students with very strong mathematical background).

Another potential use of the tutorial is to expose students to the
scientific computing pipeline *before* they have multi-variable
calculus, linear algebra, and differential equations under their belt.
I don't know of other texts with ODE models that address such an audience.

=== Referee ===

By setting the bar so low, the text will bore to death most students,
or just persuade them that the subject is not stretching them
enough. If all they need is to solve a numerical problem without
bothering about the underlying numerics, they may use a library
routine and Numerical Recipes. My assumption is that any lecture
course in scientific computing must actually treat students as adults,
stretch them and use their intelligence. I honestly don't believe that
the first part of the book is doing this. Moreover, I am quite sure
that the author, who is a very knowledgeable scientific computing
expert, is capable of writing a much more suitable text.

=== Response ===

I agree with the referee when it comes to students who specialize in
mathematical topics, but this text has a much broader audience in
sight. The referee points to an audience who will jump over the math
part, basically because they know all this (perhaps not the numerical
artifacts in detail), but the programming topics are much less known,
so this part is useful.

It might be interesting to add that critical comments similar to the
referee's have been issued regarding the software engineering part. It
is trivial, shallow, and boring - and of much lower quality than the
text in true theoretical software engineering books. At the University
of Oslo, even the beginning course in programming (based on TCSE6) in
the first semester go
further than this tutorial. But again, lots of students do not have
any background on software engineering, and they appreciate the more
to-the-point and compact Chapter 5 rather than thick software
engineering books with much theory, little practice, and ``high
level''.  The similarity to the comments on the mathematical content
of the tutorial is striking...

About 150 students have been required to do Project 4.15 in this
tutorial as a
compulsory project at the beginning of the mentioned course.  Although
many find the mathematics of this problem extremely simple, it is
disappointing to see how few who are capable of writing a clean
program with good tests for this problem. The surprisingly low
performance in the first class of students led to much of the writing
of the programming details in this tutorial. Even those who have taken
a course based on TCSE6, where all these programming topics are dealt
with in detail, need to refresh software engineering topics since
hardly any course on mathematical subjects pay attention to this
(essential) part of computing. On the other hand, any student seems
capable of making a functioning program for $u'=-u$; it is the
real-world version of this problem, with lots of parameters, that
quickly leads to spaghetti code with bugs.

TCSE1 is a book more aligned with the referee's wishes. However, after
having using that book for over 10 years in the mentioned course, it
was a common feedback from students that far too many details were
left out in the text. It certainly stretched the students, but it
didn't get them sufficiently excited to experiment with the theory
through codes for hours in the evenings. Given that the software
content in TCSE1 is also outdated, I decided to write up the numerics and
software topics for PDEs again, but this time with all nuts and bolts
visible. It has been a much greater success as teaching material,
based on student feedback and reduced drop-out in the course.

=== Referee ===

(Incidentally, this attempt to infantilise the subject often
leads to infelicities and rather dicey statements: sometimes an
oversimplification is just misleading.)

=== Response ===

I know that this comment is subject to debate when it comes to the
final chapter, but I do hope there are no serious ``infelicities and
rather dicey statements'' in the mathematical parts.

=== Referee ===

However, I loved the second part of the book, on the programming and
software engineering side of things. It is replete with useful
material, moreover much of this material is not accessible in standard
texts.

=== Response ===

The software engineering topics are important for high-quality, reliable
scientific software, but none of the numerous books on ``numerical methods
with Matlab'' that I have examined contain any information on such topics.
On the other hand, experts in software engineering will most likely
claim that the level in Chapter 5 should be much higher...

=== Referee ===

Of course, in this moment of time, an obvious objection to my opinion
is that ``well, the guy is clearly a numerical analyst who knows
mathematics but not software engineering, hence the first part of the
book is boring and the second is interesting''. Guilty as charged,
however I believe that I am quite typical. Who is studying scientific
computing, in greater generality who needs scientific computing?
Neither mathematicians nor computer scientists. Mathematicians might
be interested in computational mathematics which is a different kettle
of fish), computer scientists in, well, computer science and a
discrete world. Typical audience for this material are budding
scientists and engineers, and those will have a great deal of
mathematical foundation - but none in software engineering - in the
course of their studies.

=== Response ===

I can add that we have quite some students from computer science, with
a relatively weak mathematical background, who go for PDE-based
simulation in a master's thesis. These students are in fact capable of
learning the minimum of mathematics to enter such an application
domain. The result is often amazing software and visualizations, and
then they can be hired as scientific programmers. But this audience
needs gentle mathematics to get motivated. And
students from geoscience and bioscience are not mentally tuned to be
stretched by mathematics. I simply try to reach a different audience
than the one the referee is thinking of.

=== Referee ===

Hence my recommendation is to toughen up the first part, make it more
enticing mathematically, to push the students harder, while retaining
the second part.

=== Response ===

This is a very valid recommendation, but I think the unique selling point
of the book is the soft and accessible treatment of both parts.