### 2015-10-10 doc/src/chapters/genz/slides_genz-beamer.tex

% LaTeX Beamer file automatically generated from DocOnce
% https://github.com/hplgit/doconce

%-------------------- begin beamer-specific preamble ----------------------

\documentclass{beamer}

\usetheme{red_shadow}
\usecolortheme{default}

% turn off the almost invisible, yet disturbing, navigation symbols:
\setbeamertemplate{navigation symbols}{}

% Examples on customization:
%\usecolortheme[named=RawSienna]{structure}
%\usetheme[height=7mm]{Rochester}
%\setbeamerfont{frametitle}{family=\rmfamily,shape=\itshape}
%\setbeamertemplate{items}[ball]
%\setbeamertemplate{blocks}[rounded][shadow=true]
%\useoutertheme{infolines}
%
%\usefonttheme{}
%\useinntertheme{}
%
%\setbeameroption{show notes}
%\setbeameroption{show notes on second screen=right}

% fine for B/W printing:
%\usecolortheme{seahorse}

\usepackage{pgf}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{relsize}

\usepackage{fancybox}  % make sure fancybox is loaded before fancyvrb

\usepackage{fancyvrb}
\usepackage{minted} % requires pygments and latex -shell-escape filename
%\usepackage{anslistings}
%\usepackage{listingsutf8}

\usepackage{amsmath,amssymb,bm}
%\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{colortbl}
\usepackage[english]{babel}
\usepackage{tikz}
\usepackage{framed}
% Use some nice templates
\beamertemplatetransparentcovereddynamic

% --- begin table of contents based on sections ---
% Delete this, if you do not want the table of contents to pop up at
% the beginning of each section:
% (Only section headings can enter the table of contents in Beamer
% slides generated from DocOnce source, while subsections are used
% for the title in ordinary slides.)
\AtBeginSection[]
{
  \begin{frame}<beamer>[plain]
  \frametitle{}
  %\frametitle{Outline}
  \tableofcontents[currentsection]
  \end{frame}
}
% --- end table of contents based on sections ---

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\newcommand{\shortinlinecomment}[3]{\note{\textbf{#1}: #2}}
\newcommand{\longinlinecomment}[3]{\shortinlinecomment{#1}{#2}{#3}}

\definecolor{linkcolor}{rgb}{0,0,0.4}
\hypersetup{
    colorlinks=true,
    linkcolor=linkcolor,
    urlcolor=linkcolor,
    pdfmenubar=true,
    pdftoolbar=true,
    bookmarksdepth=3
    }
\setlength{\parskip}{7pt}  % {1em}

\newenvironment{doconceexercise}{}{}
\newcounter{doconceexercisecounter}
\newenvironment{doconce:movie}{}{}
\newcounter{doconce:movie:counter}

\newcommand{\subex}[1]{\noindent\textbf{#1}}  % for subexercises: a), b), etc

%-------------------- end beamer-specific preamble ----------------------

% Add user's preamble

% insert custom LaTeX commands...

\raggedbottom
\makeindex

%-------------------- end preamble ----------------------

\begin{document}

% matching end for #ifdef PREAMBLE
% #endif

\input{newcommands_keep}

% ------------------- main content ----------------------

% ----------------- title -------------------------

\title{Study guide: Generalizations of exponential decay models}

% ----------------- author(s) -------------------------

\author{Hans Petter Langtangen\inst{1,2}}
\institute{Center for Biomedical Computing, Simula Research Laboratory\inst{1}
\and
Department of Informatics, University of Oslo\inst{2}}
% ----------------- end author(s) -------------------------

\date{Oct 10, 2015
% <optional titlepage figure>
% <optional copyright>
}

\begin{frame}[plain,fragile]
\titlepage
\end{frame}

\section{Model extensions}
\label{decay:generalizations}

\begin{frame}[plain,fragile]
\frametitle{Extension to a variable coefficient; Forward and Backward Euler}

\begin{equation}
u'(t) = -a(t)u(t),\quad t\in (0,T],\quad u(0)=I
\label{decay:problem:a}
\end{equation}

The Forward Euler scheme:

\begin{equation}
\frac{u^{n+1} - u^n}{\Delta t} = -a(t_n)u^n
\end{equation}

The Backward Euler scheme:
\begin{equation}
\frac{u^{n} - u^{n-1}}{\Delta t} = -a(t_n)u^n
\end{equation}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Extension to a variable coefficient; Crank-Nicolson}

Eevaluting $a(t_{n+\half})$ and
using an average for $u$:
\begin{equation}
\frac{u^{n+1} - u^{n}}{\Delta t} = -a(t_{n+\half})\half(u^n + u^{n+1})
\end{equation}

Using an average for $a$ and $u$:
\begin{equation}
\frac{u^{n+1} - u^{n}}{\Delta t} = -\half(a(t_n)u^n + a(t_{n+1})u^{n+1})
\end{equation}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Extension to a variable coefficient; $\theta$-rule}

The $\theta$-rule unifies the three mentioned schemes,

\begin{equation}
\frac{u^{n+1} - u^{n}}{\Delta t} = -a((1-\theta)t_n + \theta t_{n+1})((1-\theta) u^n + \theta u^{n+1})
\end{equation}
or,
\begin{equation}
\frac{u^{n+1} - u^{n}}{\Delta t} = -(1-\theta) a(t_n)u^n - \theta
a(t_{n+1})u^{n+1}
\end{equation}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Extension to a variable coefficient; operator notation}

\begin{align*}
\lbrack D^+_t u &= -au\rbrack^n,\\ 
\lbrack D^-_t u &= -au\rbrack^n,\\ 
\lbrack D_t u &= -a\overline{u}^t\rbrack^{n+\half},\\ 
\lbrack D_t u &= -\overline{au}^t\rbrack^{n+\half}\\ 
\end{align*}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Extension to a source term}

\label{decay:source}

\begin{equation}
u'(t) = -a(t)u(t) + b(t),\quad t\in (0,T],\quad u(0)=I
\label{decay:problem:ab}
\end{equation}

\begin{align*}
\lbrack D^+_t u &= -au + b\rbrack^n,\\ 
\lbrack D^-_t u &= -au + b\rbrack^n,\\ 
\lbrack D_t u   &= -a\overline{u}^t + b\rbrack^{n+\half},\\ 
\lbrack D_t u   &= \overline{-au+b}^t\rbrack^{n+\half}
\end{align*}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Implementation of the generalized model problem}

\label{decay:general}

\begin{equation}
u^{n+1} = ((1 - \Delta t(1-\theta)a^n)u^n
+ \Delta t(\theta b^{n+1} + (1-\theta)b^n))(1 + \Delta t\theta a^{n+1})^{-1}
\end{equation}

Implementation where $a(t)$ and $b(t)$ are given as
Python functions (see file \href{{http://tinyurl.com/ofkw6kc/genz/decay_vc.py}}{\nolinkurl{decay_vc.py}}):

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def solver(I, a, b, T, dt, theta):
    """
    Solve u'=-a(t)*u + b(t), u(0)=I,
    for t in (0,T] with steps of dt.
    a and b are Python functions of t.
    """
    dt = float(dt)            # avoid integer division
    Nt = int(round(T/dt))     # no of time intervals
    T = Nt*dt                 # adjust T to fit time step dt
    u = zeros(Nt+1)           # array of u[n] values
    t = linspace(0, T, Nt+1)  # time mesh

    u[0] = I                  # assign initial condition
    for n in range(0, Nt):    # n=0,1,...,Nt-1
        u[n+1] = ((1 - dt*(1-theta)*a(t[n]))*u[n] + \ 
                  dt*(theta*b(t[n+1]) + (1-theta)*b(t[n])))/\ 
                  (1 + dt*theta*a(t[n+1]))
    return u, t
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Implementations of variable coefficients; functions}

Plain functions:

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def a(t):
    return a_0 if t < tp else k*a_0

def b(t):
    return 1
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Implementations of variable coefficients; classes}

Better implementation: class with the parameters \texttt{a0}, \texttt{tp}, and \texttt{k}
as attributes and a \emph{special method} \Verb!__call__! for evaluating $a(t)$:

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
class A:
    def __init__(self, a0=1, k=2):
        self.a0, self.k = a0, k

    def __call__(self, t):
        return self.a0 if t < self.tp else self.k*self.a0

a = A(a0=2, k=1)  # a behaves as a function a(t)
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Implementations of variable coefficients; lambda function}

\index{lambda functions}

Quick writing: a one-liner \emph{lambda function}
\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
a = lambda t: a_0 if t < tp else k*a_0
\end{minted}

In general,
\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
f = lambda arg1, arg2, ...: expressin
\end{minted}
is equivalent to
\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def f(arg1, arg2, ...):
    return expression
\end{minted}

One can use lambda functions directly in calls:
\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
u, t = solver(1, lambda t: 1, lambda t: 1, T, dt, theta)
\end{minted}
for a problem $u'=-u+1$, $u(0)=1$.

A lambda function can appear anywhere where a variable can appear.
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Verification via trivial solutions}

\label{decay:verify:trivial}

\begin{itemize}
 \item Start debugging of a new code with trying a problem
   where $u=\hbox{const} \neq 0$.

 \item Choose $u=C$ (a constant). Choose any $a(t)$ and set
   $b=a(t)C$ and
   $I=C$.

 \item "All" numerical methods will reproduce $u=_{\hbox{const}}$
   exactly (machine precision).

 \item Often $u=C$ eases debugging.

 \item In this example: \emph{any error} in the formula for $u^{n+1}$
   make $u\neq C$!
\end{itemize}

\noindent
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Verification via trivial solutions; test function}

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def test_constant_solution():
    """
    Test problem where u=u_const is the exact solution, to be
    reproduced (to machine precision) by any relevant method.
    """
    def u_exact(t):
        return u_const

    def a(t):
        return 2.5*(1+t**3)  # can be arbitrary

    def b(t):
        return a(t)*u_const

    u_const = 2.15
    theta = 0.4; I = u_const; dt = 4
    Nt = 4  # enough with a few steps
    u, t = solver(I=I, a=a, b=b, T=Nt*dt, dt=dt, theta=theta)
    print u
    u_e = u_exact(t)
    difference = abs(u_e - u).max()  # max deviation
    tol = 1E-14
    assert difference < tol
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Verification via manufactured solutions}

\label{decay:MMS}

\index{method of manufactured solutions}
\index{MMS (method of manufactured solutions)}

\begin{itemize}
 \item Choose \emph{any} formula for $u(t)$

 \item Fit $I$, $a(t)$, and $b(t)$ in $u'=-au+b$, $u(0)=I$,
   to make the chosen formula a solution of the ODE problem

 \item Then we can always have an analytical solution (!)

 \item Ideal for verification: testing convergence rates

 \item Called the \emph{method of manufactured solutions} (MMS)

 \item Special case: $u$ linear in $t$, because all sound numerical
   methods will reproduce a linear $u$ exactly (machine precision)

 \item $u(t) = ct + d$. $u(0)=0$ means $d=I$

 \item ODE implies $c = -a(t)u + b(t)$

 \item Choose $a(t)$ and $c$, and set $b(t) = c + a(t)(ct + I)$

 \item Any error in the formula for $u^{n+1}$ makes $u\neq ct+I$!
\end{itemize}

\noindent
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Linear manufactured solution}

$u^n = ct_n+I$ fulfills the discrete
equations!

First,
\begin{align}
\lbrack D_t^+ t\rbrack^n &= \frac{t_{n+1}-t_n}{\Delta t}=1,
\label{decay:fd2:Dop:tn:fw}\\ 
\lbrack D_t^- t\rbrack^n &= \frac{t_{n}-t_{n-1}}{\Delta t}=1,
\label{decay:fd2:Dop:tn:bw}\\ 
\lbrack D_t t\rbrack^n &= \frac{t_{n+\half}-t_{n-\half}}{\Delta t}=\frac{(n+\half)\Delta t - (n-\half)\Delta t}{\Delta t}=1\label{decay:fd2:Dop:tn:cn}
\end{align}

Forward Euler:

\[ [D^+ u = -au + b]^n \]

$a^n=a(t_n)$, $b^n=c + a(t_n)(ct_n + I)$, and $u^n=ct_n + I$
results in

\[ c = -a(t_n)(ct_n+I) + c + a(t_n)(ct_n + I) = c \]
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Test function for linear manufactured solution}

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def test_linear_solution():
    """
    Test problem where u=c*t+I is the exact solution, to be
    reproduced (to machine precision) by any relevant method.
    """
    def u_exact(t):
        return c*t + I

    def a(t):
        return t**0.5  # can be arbitrary

    def b(t):
        return c + a(t)*u_exact(t)

    theta = 0.4; I = 0.1; dt = 0.1; c = -0.5
    T = 4
    Nt = int(T/dt)  # no of steps
    u, t = solver(I=I, a=a, b=b, T=Nt*dt, dt=dt, theta=theta)
    u_e = u_exact(t)
    difference = abs(u_e - u).max()  # max deviation
    print difference
    tol = 1E-14  # depends on c!
    assert difference < tol
\end{minted}
\end{frame}

\section{Computing convergence rates}

\begin{frame}[plain,fragile]
\frametitle{Computing convergence rates}

\label{decay:convrates}

Frequent assumption on the relation between the numerical error $E$ and
some discretization parameter $\Delta t$:

\begin{equation}
E = C\Delta t^r,
\label{decay:E:dt}
\end{equation}

\begin{itemize}
 \item Unknown: $C$ and $r$.

 \item Goal: estimate $r$ (and $C$) from numerical experiments
\end{itemize}

\noindent

\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Estimating the convergence rate $r$}

Perform numerical experiments: $(\Delta t_i, E_i)$, $i=0,\ldots,m-1$.
Two methods for finding $r$ (and $C$):

\begin{enumerate}
 \item Take the logarithm of (\ref{decay:E:dt}), $\ln E = r\ln \Delta t + \ln C$,
    and fit a straight line to the data points $(\Delta t_i, E_i)$,
    $i=0,\ldots,m-1$.

 \item Consider two consecutive experiments, $(\Delta t_i, E_i)$ and
    $(\Delta t_{i-1}, E_{i-1})$. Dividing the equation
    $E_{i-1}=C\Delta t_{i-1}^r$ by $E_{i}=C\Delta t_{i}^r$ and solving
    for $r$ yields
\end{enumerate}

\noindent
\begin{equation}
r_{i-1} = \frac{\ln (E_{i-1}/E_i)}{\ln (\Delta t_{i-1}/\Delta t_i)}
\label{decay:conv:rate}
\end{equation}
for $i=1,=\ldots,m-1$.

Method 2 is best.
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Brief implementation}

Compute $r_0, r_1, \ldots, r_{m-2}$ from $E_i$ and $\Delta t_i$:

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def compute_rates(dt_values, E_values):
    m = len(dt_values)
    r = [log(E_values[i-1]/E_values[i])/
         log(dt_values[i-1]/dt_values[i])
         for i in range(1, m, 1)]
    # Round to two decimals
    r = [round(r_, 2) for r_ in r]
    return r
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{We embed the code in a real test function}

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
def test_convergence_rates():
    # Create a manufactured solution
    # define u_exact(t), a(t), b(t)

    dt_values = [0.1*2**(-i) for i in range(7)]
    I = u_exact(0)

    for theta in (0, 1, 0.5):
        E_values = []
        for dt in dt_values:
            u, t = solver(I=I, a=a, b=b, T=6, dt=dt, theta=theta)
            u_e = u_exact(t)
            e = u_e - u
            E = sqrt(dt*sum(e**2))
            E_values.append(E)
        r = compute_rates(dt_values, E_values)
        print 'theta=%g, r: %s' % (theta, r)
        expected_rate = 2 if theta == 0.5 else 1
        tol = 0.1
        diff = abs(expected_rate - r[-1])
        assert diff < tol
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{The manufactured solution can be computed by sympy}

We choose $\uex(t) = \sin(t)e^{-2t}$, $a(t)=t^2$, fit $b(t)=u'(t)-a(t)$:

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
# Create a manufactured solution with sympy
import sympy as sym
t = sym.symbols('t')
u_exact = sym.sin(t)*sym.exp(-2*t)
a = t**2
b = sym.diff(u_exact, t) + a*u_exact

# Turn sympy expressions into Python function
u_exact = sym.lambdify([t], u_exact, modules='numpy')
a = sym.lambdify([t], a, modules='numpy')
b = sym.lambdify([t], b, modules='numpy')
\end{minted}

Complete code: \href{{http://tinyurl.com/ofkw6kc/genz/decay_vc.py}}{\nolinkurl{decay_vc.py}}.
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Execution}

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{console}
Terminal> python decay_vc.py
...
theta=0, r: [1.06, 1.03, 1.01, 1.01, 1.0, 1.0]
theta=1, r: [0.94, 0.97, 0.99, 0.99, 1.0, 1.0]
theta=0.5, r: [2.0, 2.0, 2.0, 2.0, 2.0, 2.0]
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Debugging via convergence rates}

Potential bug: missing \texttt{a} in the denominator,

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
u[n+1] = (1 - (1-theta)*a*dt)/(1 + theta*dt)*u[n]
\end{minted}
Running \Verb!decay_convrate.py! gives same rates.

Why? The value of $a$... ($a=1$)

0 and 1 are \emph{bad values} in tests!

Better:
\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{console}
Terminal> python decay_convrate.py --a 2.1 --I 0.1  \ 
          --dt 0.5 0.25 0.1 0.05 0.025 0.01
...
Pairwise convergence rates for theta=0:
1.49 1.18 1.07 1.04 1.02

Pairwise convergence rates for theta=0.5:
-1.42 -0.22 -0.07 -0.03 -0.01

Pairwise convergence rates for theta=1:
0.21 0.12 0.06 0.03 0.01
\end{minted}

Forward Euler works...because $\theta=0$ hides the bug.

This bug gives $r\approx 0$:

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
u[n+1] = ((1-theta)*a*dt)/(1 + theta*dt*a)*u[n]
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Extension to systems of ODEs}

Sample system:

\begin{align}
u' &= a u + bv\\ 
v' &= cu +  dv
\end{align}

The Forward Euler method:

\begin{align}
u^{n+1} &= u^n + \Delta t (a u^n + b v^n)\\ 
v^{n+1} &= u^n + \Delta t (cu^n + dv^n)
\end{align}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{The Backward Euler method gives a system of algebraic equations}

The Backward Euler scheme:

\begin{align}
u^{n+1} &= u^n + \Delta t (a u^{n+1} + b v^{n+1})\\ 
v^{n+1} &= v^n + \Delta t (c u^{n+1} + d v^{n+1})
\end{align}
which is a $2\times 2$ linear system:

\begin{align}
(1 - \Delta t a)u^{n+1} + bv^{n+1} &= u^n \\ 
c u^{n+1} + (1 - \Delta t d) v^{n+1} &= v^n
\end{align}

Crank-Nicolson also gives a $2\times 2$ linear system.
\end{frame}

\section{Methods for general first-order ODEs}
\label{decay:1stODEs}

\begin{frame}[plain,fragile]
\frametitle{Generic form}

The standard form for ODEs:
\begin{equation}
u' = f(u,t),\quad u(0)=I
\label{decay:ode:general}
\end{equation}

$u$ and $f$: scalar or vector.

Vectors in case of ODE systems:
\[ u(t) = (u^{(0)}(t),u^{(1)}(t),\ldots,u^{(m-1)}(t))   \]

\begin{align*}
f(u, t) = ( & f^{(0)}(u^{(0)},\ldots,u^{(m-1)})\\ 
            & f^{(1)}(u^{(0)},\ldots,u^{(m-1)}),\\ 
            & \vdots\\ 
            & f^{(m-1)}(u^{(0)}(t),\ldots,u^{(m-1)}(t)))
\end{align*}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{The $\theta$-rule}

\begin{equation}
\frac{u^{n+1}-u^n}{\Delta t} = \theta f(u^{n+1},t_{n+1}) +
(1-\theta)f(u^n, t_n)
\label{decay:fd2:theta}
\end{equation}
Bringing the unknown $u^{n+1}$ to the left-hand side and the known terms
on the right-hand side gives

\index{implicit schemes} \index{explicit schemes} \index{theta-rule} \index{$\theta$-rule}

\begin{equation}
u^{n+1} - \Delta t \theta f(u^{n+1},t_{n+1}) =
u^n + \Delta t(1-\theta)f(u^n, t_n)
\end{equation}
This is a \emph{nonlinear} equation in $u^{n+1}$ (unless $f$ is linear in $u$)!
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Implicit 2-step backward scheme}

\index{backward scheme, 2-step} \index{BDF2 scheme}

\[ u'(t_{n+1}) \approx \frac{3u^{n+1} - 4u^{n} + u^{n-1}}{2\Delta t}\]

Scheme:
\[ u^{n+1} = \frac{4}{3}u^n - \frac{1}{3}u^{n-1} +
\frac{2}{3}\Delta t f(u^{n+1}, t_{n+1})
\label{decay:fd2:bw:2step}
\]
Nonlinear equation for $u^{n+1}$.
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{The Leapfrog scheme}

\index{Leapfrog scheme}

Idea:
\begin{equation}
u'(t_n)\approx \frac{u^{n+1}-u^{n-1}}{2\Delta t} = [D_{2t} u]^n
\end{equation}

Scheme:

\[ [D_{2t} u = f(u,t)]^n\]
or written out,
\begin{equation}
u^{n+1} = u^{n-1} + \Delta t f(u^n, t_n)
\label{decay:fd2:leapfrog}
\end{equation}

\begin{itemize}
 \item Some other scheme must be used as starter ($u^1$).

 \item Explicit scheme - a nonlinear $f$ (in $u$) is trivial to handle.

 \item Downside: Leapfrog is always unstable after some time.
\end{itemize}

\noindent
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{The filtered Leapfrog scheme}

\index{Leapfrog scheme, filtered}

After computing $u^{n+1}$, stabilize Leapfrog by
\begin{equation}
u^n\ \leftarrow\ u^n + \gamma (u^{n-1} - 2u^n + u^{n+1})
\label{decay:fd2:leapfrog:filtered}
\end{equation}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{2nd-order Runge-Kutta scheme}

\index{Heun's method}
\index{Runge-Kutta, 2nd-order scheme}

Forward-Euler + approximate Crank-Nicolson:
\begin{align}
u^* &= u^n + \Delta t f(u^n, t_n),
\label{decay:fd2:RK2:s1}\\ 
u^{n+1} &= u^n + \Delta t \half \left( f(u^n, t_n) + f(u^*, t_{n+1})
\right)
\label{decay:fd2:RK2:s2}
\end{align}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{4th-order Runge-Kutta scheme}

\label{decay:fd2:RK4}

\begin{itemize}
 \item The most famous and widely used ODE method

 \item 4 evaluations of $f$ per time step

 \item Its \href{{http://tinyurl.com/nclmcng/pub/sphinx-decay/._main_decay007.html#th-order-runge-kutta-scheme}}{derivation} is a very good illustration of numerical thinking!
\end{itemize}

\noindent
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{2nd-order Adams-Bashforth scheme}

\index{Adams-Bashforth scheme, 2nd order}

\begin{equation}
u^{n+1} = u^n + \half\Delta t\left( 3f(u^n, t_n) - f(u^{n-1}, t_{n-1})
\right)
\label{decay:fd2:AB2}
\end{equation}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{3rd-order Adams-Bashforth scheme}

\index{Adams-Bashforth scheme, 3rd order}

\begin{equation}
u^{n+1} = u^n + \frac{1}{12}\left( 23f(u^n, t_n) - 16 f(u^{n-1},t_{n-1})
+ 5f(u^{n-2}, t_{n-2})\right)
\label{decay:fd2:AB3}
\end{equation}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{The Odespy software}

\href{{https://github.com/hplgit/odespy}}{Odespy}
features simple Python implementations of the most fundamental
schemes as well as Python interfaces to several famous packages for
solving ODEs: \href{{https://computation.llnl.gov/casc/odepack/odepack_home.html}}{ODEPACK},
\href{{https://computation.llnl.gov/casc/odepack/odepack_home.html}}{Vode},
\href{{http://www.netlib.org/ode/rkc.f}}{rkc.f},
\href{{http://www.netlib.org/ode/rkf45.f}}{rkf45.f},
\href{{http://www.unige.ch/~hairer/software.html}}{Radau5}, as well
as the ODE solvers in
\href{{http://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.ode.html}}{SciPy},
\href{{http://docs.sympy.org/dev/modules/mpmath/calculus/odes.html}}{SymPy}, and
\href{{http://olivierverdier.github.com/odelab/}}{odelab}.

Typical usage:

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{text}
# Define right-hand side of ODE
def f(u, t):
    return -a*u

import odespy
import numpy as np

# Set parameters and time mesh
I = 1; a = 2; T = 6; dt = 1.0
Nt = int(round(T/dt))
t_mesh = np.linspace(0, T, Nt+1)

# Use a 4th-order Runge-Kutta method
solver = odespy.RK4(f)
solver.set_initial_condition(I)
u, t = solver.solve(t_mesh)
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Example: Runge-Kutta methods}

\begin{minted}[fontsize=\fontsize{9pt}{9pt},linenos=false,mathescape,baselinestretch=1.0,fontfamily=tt,xleftmargin=2mm]{python}
solvers = [odespy.RK2(f),
           odespy.RK3(f),
           odespy.RK4(f),
           odespy.BackwardEuler(f, nonlinear_solver='Newton')]

for solver in solvers:
    solver.set_initial_condition(I)
    u, t = solver.solve(t)

# + lots of plot code...
\end{minted}
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Plots from the experiments}

% inline figure
\centerline{\includegraphics[width=0.9\linewidth]{fig-genz/decay_odespy1_png.png}}

The 4-th order Runge-Kutta method (\texttt{RK4}) is the method of choice!
\end{frame}

\begin{frame}[plain,fragile]
\frametitle{Example: Adaptive Runge-Kutta methods}

\begin{itemize}
 \item Adaptive methods find "optimal" locations of the mesh points
   to ensure that the error is less than a given tolerance.

 \item Downside: approximate error estimation, not always optimal
   location of points.

 \item "Industry standard ODE solver": Dormand-Prince 4/5-th order
   Runge-Kutta (MATLAB's famous \texttt{ode45}).
\end{itemize}

\noindent
% inline figure
\centerline{\includegraphics[width=0.9\linewidth]{fig-genz/decay_DormandPrince_adaptivity.png}}
\end{frame}

\end{document}